---
title: "Reproducible Research project 1"
output:
  word_document: default
  html_document: default
---
Reading the file

```{r, echo= TRUE}
filedata <- read.csv("activity.csv")
names(filedata)
head(filedata)
```

Calculating total number of steps per day:
```{r, echo=TRUE}
stepsperday <- aggregate(steps~date, filedata, sum, na.rm= TRUE)
hist(stepsperday$steps, main = "Steps taken per day", xlab = "stepsperday", ylab = "steps")
```

Calculating mean and median total number of steps per day:
```{r, echo=TRUE}
meanstepstakenperday <- mean(stepsperday$steps)
meanstepstakenperday
medianstepstaken <- median(stepsperday$steps)
medianstepstaken                           
```

Calculating average number of steps taken across all days:
```{r, echo=TRUE}
everyinterval<- aggregate(steps~interval, data=filedata, mean, na.rm=TRUE)
plot(steps~interval, data=everyinterval, type="l")
```

Maximum number of steps:
```{r, echo=TRUE}
maxsteps <- everyinterval[which.max(everyinterval$steps),]$interval
maxsteps
```

Calculating missing values:
```{r, echo=TRUE}
missingval <- sum(is.na(filedata$steps))
missingval
meaneveryinterval<-function(interval){
    everyinterval[everyinterval$interval==interval,]$steps
}
```

Creating new dataset with missing values:
```{r, echo=TRUE}
newdataset <- filedata
for(i in 1:nrow(newdataset)){
    if(is.na(newdataset[i,]$steps)){
        newdataset[i,]$steps <- meaneveryinterval(newdataset[i,]$interval)
    }
}
stepsperdaynew <- aggregate(steps ~ date, data=newdataset, sum)
hist(stepsperdaynew$steps, main = "Total steps per day")
```

mean and median of new datset: 
```{r, echo=TRUE}
meannew <- mean(stepsperdaynew$steps)
meannew
mediannew <- median(stepsperdaynew$steps)
```

Indicating whether a given date is a weekday or weekend:
```{r, echo=TRUE}
newdataset$date <- as.Date(strptime(newdataset$date, format="%Y-%m-%d"))
newdataset$day <- weekdays(newdataset$date)
for (i in 1:nrow(newdataset)) {
    if (newdataset[i,]$day %in% c("Saturday","Sunday")) {
        newdataset[i,]$day<-"weekend"
    }
    else{
        newdataset[i,]$day<-"weekday"
    }
}
stepsbyday <- aggregate(newdataset$steps ~ newdataset$interval + newdataset$day, newdataset, mean)
```

Plotting the average number of steps taken on x-axis and averaged across all weekday days or weekend days on y-axis:
```{r, echo=TRUE}
names(stepsbyday) <- c("interval", "day", "steps")
library(lattice)
xyplot(steps ~ interval | day, stepsbyday, type = "l", layout = c(1, 2), 
    xlab = "Interval", ylab = "Number of steps")
```










